---
title: "Getting started with R"
output: html_notebook
---

This notebook contains the answers to Task 1: Get started with R.

This notebook is intended to learn the basics of R.

```{r - Initial imports}
library(readr)

```


```{r - Read the dataset from memory}
df_car<- read.csv("../cars.csv")

```

*Understand the data:*

```{r - Getting to know the data}
print("List your attributes within your data set.")
attributes(df_car)
print("Prints the min, max, mean, median, and quartiles of each attribute.")
summary(df_car)
print("Displays the structure of your data set.")
str(df_car)
print("Names your attributes within your data set.")
names(df_car)
print("Will print out the instances within that particular column in your data set.")
df_car$speed.of.car #
df_car$distance.of.car #Will print out the instances within that particular column in your data set.
```

```{r - Do column types coincide with their data? Or do I need to convert data types?}
sapply(df_car, typeof)

#df_car$ColumnName<-integer(df_car$ColumnName)



```

```{r - Rename columns}
names(df_car)<-c("CarMaker","Speed","Dist") 

```


```{r - Are there missing values?}
summary(df_car)

any(is.na(df_car))

```
There are no missing values in the car_df, thus the chunk below is unnecessary.

```{r - Address missing values}
#na.omit(DatasetName$ColumnName#Drops any rows with missing values and omits them forever.
#na.exclude(DatasetName$ColumnName)#Drops any rows with missing values, but keeps track of where they were.
#DatasetName$ColumnName[is.na(DatasetName$ColumnName)]<-mean(DatasetName$ColumnName,na.rm = TRUE) #Replace the missing values with the mean
```

*Exploratory Data Analysis*

```{r - Histogram plots}
hist(df_car$Speed)
hist(df_car$Dist)
```



```{r - Scatter (Box) Plot}
plot(df_car$Speed,df_car$Dist)
```


```{r - Normal Quantile Plot}
qqnorm(df_car$Speed)
qqnorm(df_car$Dist)
```

As can be seen above, the relationship between distance and speed is pretty close to a line, thus a linear regression will probably yield good results when modelling one of these variables.

*Creating testing and training sets:*



```{r - Set seed and find row length for train and test datasets}
set.seed(42)

trainSize<-round(nrow(df_car)*0.7) #70% for train dataset
testSize<-nrow(df_car)-trainSize #Remaining 30% for test dataset


trainSize
testSize
```
```{r - Create test and train datasets}
training_indices<-sample(seq_len(nrow(df_car)),size =trainSize) # Select random sample of row numbers with size 'trainsize'

trainSet<-df_car[training_indices,] #TrainSet will contain the selected indices.

testSet<-df_car[-training_indices,] #TestSet will contain only the remaining rows that have not been selected.
```

*Linear Regression Model*

```{r - Create model}
DistModel<-lm(Dist ~ Speed-1, trainSet) # response(y) ~ terms(X)
# used -1 to omit intercept
summary(DistModel)
```
The multiple R-squared is 0.943 which is preety close to 1, this models seems to be a good fit.

The p-value is clearly smaller than 0.05 meaning that the relationship between these two variables is statistically significant.

Adjusted R-squared is used to determine how reliable the correlation is and how much it is determined by the addition of independent variables.

F-statistics are based on the ratio of mean squares. The term “mean squares” may sound confusing but it is simply an estimate of population variance that accounts for the degrees of freedom (DF) used to calculate that estimate.

```{r - Predictions}
DistPredictions <- predict(DistModel,testSet)
cbind(Pred_Dist = DistPredictions,Real_Dist = testSet$Dist)
```

